Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-04-03 paper/src/paper.tex
paper fix

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing (HPC) platforms. Despite
  its good properties (simplicity, no starvation), it could still be improved
  on a per-system basis. This tuning process is difficult because of
  non-linearities in the scheduling process. The study proposed here considers
  an online approach to the automatic tuning of the EASY heuristic for HPC
  platforms. More precisely, we consider the problem of selecting a reordering
  policy of the job queue under several feedback modes. We show via a
  comprehensive experimental campaign that noisy feedback (using a weak
  simulator) recovers existing in-hindsight results that allow to divide the
  average waiting time up to a factor of 2. Moreover, we show that bandit
  feedback can be used by a simple multi-armed bandit algorithm to decrease the
  average waiting time down to 40\% of its original value without using a
  simulator.

@@

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of modern society (including climate change,
health, green energy or security) is a strategic challenge. The main pillar of
the answer to this challenge is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity does
not only lead to sophisticated specific Exascale platforms, but the objective
of the community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly, system
management software still has a long way to go. The existing job and resource
management softwares allow to run tens of thousands of jobs on hundreds of thousands
of cores. They are based on robust policies, however, despite their positive
properties such as relative simplicity and prevention of job starvation,
there is still room for improvement. We propose here to study how to tune the
ordering of the submission queues in the frame of the classical
EASY-BackFilling family of heuristics using two new online learning approaches
(an approach based on a simulator and an approach using multi-armed bandit).

@@

Contrarily to most existing studies, which consider the learning of specific
parameters of a trace, we provide a complete system that uses feedback from the
system for adapting the whole behavior online.  Our focus within this work is
to learn how the management system behaves within the usual framework of
EASY-BF. More precisely, we provide strategies for choosing the most
appropriate queue reordering policy to be used for job selection. This approach
is assessed by an extensive resampling-based experimental campaign that uses 7
actual traces.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. We focus in this work on the
average waiting time of the jobs. Each method aims to choose the best queue
reordering policy in a fixed search set of semantically diverse options.

@@

The second method (bandit feedback) is a simulator-less approach based on a
multi-armed bandit algorithm that derives feedback by measuring system
performance.

Both of these methods are evaluated with a specially developped simulator that is
open-sourced and packaged for the community.

Our main results are as follows: The approach based on the simulator provides
very good results, close to the best policy in the search set. In essence, this
means that we are able to recover in an online fashion existing results that
reduce average waiting times between -11\% and -60\% compared to EASY-FCFS.
Indeed, we show that the required sample sizes for policy selection in batch
schedulers are small enough for workload resampling to be avoided in practice.
The variant based on noisy data, which is closer to actual conditions of
uncertainties on both jobs and platform, also behaves very well. The alternate
approach based on a multi-armed bandit also provides reasonable results at a
much lower cost, reducing the average waiting times down to between -8\% and
-46\% compared to EASY-FCFS.

Let us emphasize the main results of this study: While the First-Come,
First-Serve approach to minimizing the waiting times is a common default
strategy, it can absolutely be outperformed.  We show how to automate selection
of a good alternative policy in two cases (depending whether a simulator is
available or not), and provide the difference in performance to be expected
between both situations.

@@

This section presents and discuss the most significant results related to job
scheduling and learning algorithms in this context.  Let us start by recalling
some basics about scheduling heuristics in HPC platforms:

Parallel job scheduling is a old studied theoretical
problem~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865}
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied on the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead, which mainly explains why most of the
machines of the top500 ranking~\cite{top500} rather use a variant of EASY-BF
instead.  There is a large body of work seeking to improve/tune EASY. Indeed,
while the basic mechanism is used by some actual resource and job management
softwares (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
without fine tunings by system administrators.

@@

The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
method to tuning these queues, although it requires simulated scheduling runs
at decision time and therefore, it costs much more than the natural execution of EASY.

@@

Several works use similar techniques in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.  However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
which directly learn a good scheduling policy from a given policy space.

Existing work~\cite{jsspp17} takes this approach in an offline manner, by
splitting workload data in a training/testing fashion. This work studied
EASY-Backfilling and shows that the relative performance of some well-known
priority orders for starting jobs differs between workloads, but is relatively
stable troughout time for a given workload.

@@

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (represented by an arm
of the Bandit) must be chosen in a fixed set and the corresponding reward is
observed. The goal of a MAB algorithm is to maximize the total reward obtained
in a successive number of rounds.  There exist a bunch of works that address
this problem under a variety of constraints, The two most popular settings are
the original stochastic  case and the adversarial case. See~\cite{thompson} for
the original work on the stochastic case and~\cite{Auer2002} for the "Upper
Confidence Bound" family of algorithms. According to our knowledge, the work
of~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
to the review~\cite{bubnow} for a comprehensive overview of the field. While
these algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraints such as the fact that the rewards
should be contained in a range. A simple heuristic called
\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to achieve good
results in practice, and does not have this requirement.

@@

The selection of the job to run is performed according to a scheduling policy
that establishes the order in which the jobs are executed.  EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e. every job will
be scheduled at some moment).

The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
time that requires a scheduling decision (i.e. job submission or termination),
the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
and starts jobs until it finds a job that can not be started right away. It
then makes a reservation for this job at the earliest predictable time and
starts \textit{backfilling} the job queue in FCFS order, starting any job that
does not delay the unique reservation.

@@

We present here the strategies we have retained for selecting a policy. We will
refer to the period during which a selected policy is applied as the
\textit{policy period} and will denote the length of this period as $\Delta$.
The time interval is thus divided into periods of equal lengths ($\Delta$):
$\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy
period. A new policy is selected at the beginning of each period and applied
during the whole period.

We further assume that there is a certain regularity among periods,
\textit{i.e.} that the distributions underlying the jobs submitted do not
radically differ between consecutive periods. This assumption is validated in
the study presented in~\cite{jsspp17}. It further entails that the behavior of
a policy on the previous periods reflects its behavior on the current one, so
that the selection of a policy can be based on its behavior on previous
periods. However, there may be some variability between different periods for
certain cost metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:mosn} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all the 7
policies considered. This indicates that when when the cost metric is averaged
over different periods, there is a tradeoff to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

@@

We thus explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to tradeoff
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

@@

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies
remain comparable over time, as not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

@@

This section compares the different approaches to policy selection via a
comprehensive experimental campaign.  Subsection~\ref{sub:protocol} outlines
the experimental protocol used and Subsection~\ref{sub:results} contains the
experimental results and their discussion.

@@

The Evaluation of computer systems performance is a complex
task. Here, the two main difficulties are choosing a simulation
model and taking into account the variability in the average waiting times.
This section presents our approach, which is based on lightweight simulation
and trace resampling.

@@

The choice of a simulator is a critical part of experimental validation that
raises a tradeoff between precision and runtime. High precision can be obtained
by carefully modeling the platform and its network topology, or extracting
information from the jobs from their sources or post-morted logs. This is the
approach used by high-fidelity batch scheduling simulators such as
Batsim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without adressing the allocation problem.
Moreover, the experimental protocol used here requires many simulation runs.
Therefore, we set the precision/runtime tradeoff at the point which minimizes
simulation runtimes. We discard all topological information relative to the
platform and use the job processing times of the original workloads. In this
setup, the processors are considered to be undistinguishable from each other,
and jobs can be discontinuously mapped to any available processor on the
system. We develop a lightweight simulator\cite{ocst} and release it to the
community. See the reproducibility paragraph below for more information.  With
this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in 33
seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a point of comparison, the Batsim simulator with no network
communication modeling and no resource contiguity takes more than 7 hours to
replay the same trace on the same hardware. The complete simulation campaign
presented below is executed on a single Dell PowerEdge R730 in 22 hours,
including input preparation and analysis code.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. This variability can be so extreme
that even a timeframe of a fews years can be insufficient to compare scheduling
algorithms. In our experience inverting some months in the trace at random can
lead to reversal in the results. This motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: For evey week in
a resampled trace, we iterate over every existing user in the original trace.
For each of these users, we choose a week of the original trace uniformly at
random. The jobs from this user in this original week are chosen to be present
in the week being resampled.  This simple mechanism enables us to preserve the
dependency within a week.  Although the system state may not be independent
from one week to an other, we rely on existing work~\cite{jsspp17} and assume
for these experiments that the system is in a stationary state. While we
do provide guidelines as how to incorporate tracking variant of our algorithms
in~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
this will not be studied in the experiments from this paper, and is left for
future research.

@@

We use devise the following experimental campaign. For each of the 7 workloads
considered, we use the original data to resample 60 traces, each of a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours, a sane default. The choice of this default
stems from the analysis developped in ~\cite{jsspp17}. This value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

  \item The simulation-based ("full feedback") strategy with hyperparameter
    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This choice of
    hyperparameter is a consequence of the natural rythm of human activity:
    daily or weekly simulation periods seems to facilitate comparability of
    performance between two periods. Indeed, arbitrary period values would
    needlessly increase the variability of the cost metric.

@@

The experiment campaign consists of 7 traces $\times$ 60 resamples $\times$
(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
two years.  This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times every at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

@@

We begin by attracting the attention of the reader to Table~\ref{tab:full},
which reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. More
precisely, let us denote the resampled traces of a given workload by $T_k$,
where k is the index of the resampled trace ($1 \le k \le 60$). We also denote
the number of jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by
$\bar{w}(T_k,S)$ the cumulative waiting time obtained by a simulation run for a
strategy $S$:

@@

The scheduling of parallel jobs on a given HPC platform is a very hard
optimization problem with many uncertain parameters. Even thought these
uncertainties could be reduced, determining efficient strategies remains
difficult.

@@

More precisely, we showed that it is worth to learn on the scheduling process
itself. Indeed, reordering the submission queues under EASY-BackFilling leads
to considerable gains in performance.

This approach was used in two methods: a method based on a simulator and a
method based on a multi-armed bandit algorithm. The first one provides very good results,
reducing the average waiting times of the baseline FCFS ordering policy by a
factor of -11\% to -60\%. The second one is
slighly less efficient with an improvement factor of -8\% to
-46\%. However, it is easier to use and less costly to run.

There are two directions for extensions of this work.  First, the question of
the existence of a more efficient selection strategy for bandit feedback for
these systems is not addressed by this work. Is a simulator really necessary?
%@Denis: a essayer.. 1 mois de travail pour tester?
Second, one can help but wonder what performance could be achieved by extending
the search space to arbitrary policies. This means doing a complete study
including lower bounding the cost and studying learning performance using some
function approximator.
%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
As these directions are data-bound, they will require careful application of
trace resampling techniques in order to be successful.

>>>>

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing (HPC) platforms. Despite
  its good properties (simplicity, no starvation), it could still be improved
  on a per-system basis. This tuning process is difficult because of
  non-linearities in the scheduling process. The study proposed here considers
  an online approach to the automatic tuning of the EASY heuristic for HPC
  platforms. More precisely, we consider the problem of selecting a reordering
  policy of the job queue under several feedback modes. We show via a
  comprehensive experimental campaign that noisy feedback (using a weak
  simulator) recovers existing in-hindsight results that allow to divide the
  average waiting time by almost 2. Moreover, we show that bandit feedback can
  be used by an epsilon-greedy algorithm to cut the average waiting time by
  40\% without using a simulator.

@@

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of modern society (including climate change,
health, green energy or security) is a strategic challenge. The main pillar of
the answer to this challenge is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity does
not only lead to sophisticated specific Exascale platforms, but the objective
of the community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly, system
management software still has a long way to go. The existing job and resource
management software allow to run tens of thousands of jobs on hundreds of thousands
of cores. They are based on robust policies, however, despite their positive
properties such as relative simplicity and prevention of job starvation,
there is still room for improvement. Here, we propose to study how to tune the
ordering of the submission queues in the frame of the classical
EASY-BackFilling family of heuristics using two new online learning approaches
(an approach based on a simulator and an approach using multi-armed bandit).

@@

Unlike most existing studies, which consider the learning of specific
parameters of a trace, we provide a complete system that uses feedback from the
system for adapting the whole behavior online.  Our focus within this work is
to learn how the management system behaves within the usual framework of
EASY-BF. More precisely, we provide strategies for choosing the most
appropriate queue reordering policy to be used for job selection. This approach
is assessed by an extensive resampling-based experimental campaign that uses 7
actual traces.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. In this work, we focus on the
average waiting time of the jobs. Each method aims to choose the best queue
reordering policy in a fixed search set of semantically diverse options.

@@

The second method (bandit feedback) is a simulator-less approach based on a
multi-armed bandit algorithm, which derives feedback by measuring system
performance.

Both of these methods are evaluated with a specifically developed simulator,
which is open-sourced and packaged for the community.

Our main results are as follows: The approach based on the simulator provides
very good results, close to the best policy in the search set. In essence, we
are able to recover in an online fashion existing results that reduce average
waiting times by 11\% to 60\% compared to EASY-FCFS.  Indeed, we show
that the required sample sizes for policy selection in batch schedulers are
small enough for workload resampling to be avoided in practice.  The variant
based on noisy data, which is closer to the actual conditions of uncertainties on
both jobs and platform, also behaves very well. The alternate approach based on
a multi-armed bandit also provides reasonable results at a much lower cost,
reducing the average waiting times by 8\% to 46\% compared to
EASY-FCFS.

Let us emphasize the main results of this study: While the First-Come,
First-Serve approach to minimizing the waiting times is a common default
strategy, it can absolutely be outperformed.  We show how to automate selection
of a good alternative policy in two cases (depending on whether a simulator is
available or not), and provide the difference in performance to be expected
between both situations.

@@

This section presents and discusses the most significant results related to job
scheduling and learning algorithms in this context.  Let us start by recalling
some basics about scheduling heuristics in HPC platforms:

Parallel job scheduling is a old studied theoretical
problem~\cite{Frachtenberg:2009:JSS:1692356,Feitelson:2004:PJS:2128864.2128865}
whose practical ramifications, varying hypotheses, and inherent uncertainty of
the problem applied to the HPC field have driven practitioners to use simple
heuristics (and researchers to study their behavior). The two most popular
heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead, which mainly explains why most of the
machines of the top500 ranking~\cite{top500} rather use a variant of EASY-BF
instead.  There is a large body of work seeking to improve/tune EASY. Indeed,
while the basic mechanism is used by some actual resource and job management
software (most notably SLURM~\cite{SLURMdocSCHED}), this is rarely done
without fine tuning by system administrators.

@@

The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
method to tunine these queues, although it requires simulated scheduling runs
at decision time and therefore, it costs much more than the natural execution of EASY.

@@

Several works use similar techniques in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.  However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
which directly learns a good scheduling policy from a given policy space.

Existing work~\cite{jsspp17} takes this approach in an offline manner, by
splitting workload data in a training/testing fashion. This work studies
EASY-Backfilling and shows that the relative performance of some well-known
priority orders for starting jobs differs between workloads, but is relatively
stable over time for a given workload.

@@

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (represented by an arm
of the Bandit) must be chosen in a fixed set and the corresponding reward is
observed. The goal of a MAB algorithm is to maximize the total reward obtained
in a successive number of rounds. There exist multiple works addressing
this problem under a variety of constraints, The two most popular settings are
the original stochastic  case and the adversarial case. See~\cite{thompson} for
the original work on the stochastic case and~\cite{Auer2002} for the "Upper
Confidence Bound" family of algorithms. According to our knowledge, the work
of~\cite{Banos} is the earliest for the adversarial case;
see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
to~\cite{bubnow} for a comprehensive review of the field. While
these algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraints such as the fact that the rewards
should be contained in a range. A simple heuristic called
\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to achieve good
results in practice, and does not have this requirement.

@@

The selection of the job to run is performed according to a scheduling policy,
which determines the order in which the jobs are executed. EASY-Backfilling is
the most widely used policy due to its simple and robust implementation and
known benefits such as high system utilization~\cite{easy}. This strategy has
no worst case guarantee beyond the absence of starvation (i.e. every job will
be scheduled at some moment).

The EASY-FCFS heuristic uses a job queue to select and backfill jobs.  At any
time requiring a scheduling decision (i.e. job submission or termination),
the scheduler goes through the job queue in First-Come,First-Serve (FCFS) order
and starts jobs until it finds a job that can not be started right away. Then,
it makes a reservation for this job at the earliest predictable time and
starts \textit{backfilling} the job queue in FCFS order, starting any job that
does not delay the unique reservation.

@@

Here, we present the strategies we have retained for selecting a policy. We will
refer to the period during which a selected policy is applied as the
\textit{policy period} and will denote the length of this period as $\Delta$.
The time interval is thus divided into periods of equal lengths ($\Delta$):
$\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy
period. A new policy is selected at the beginning of each period and applied
during the whole period.

We further assume that there is a certain regularity among periods,
\textit{i.e.} that the distribution of the job submission process does not
radically differ between consecutive periods. This assumption is validated in
the study presented in~\cite{jsspp17}. It further entails that the behavior of
a policy on the previous periods reflects its behavior on the current one, so
that the selection of a policy can be based on its behavior on previous
periods. However, there may be some variability between different periods for
certain cost metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:mosn} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all the 7
policies considered. This indicates that when when the cost metric is averaged
over different periods, there is a tradeoff to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

@@

Thus, we explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to tradeoff
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

@@

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
remain comparable over time, as not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

@@

This section compares the different approaches to policy selection via a
comprehensive experimental campaign.  Subsection~\ref{sub:protocol} outlines
our experimental protocol and Subsection~\ref{sub:results} contains the
experimental results and their discussion.

@@

The evaluation of computer systems performance is a complex
task. Here, the two main difficulties are choosing a simulation
model and taking into account the variability in the average waiting times.
This section presents our approach, which is based on lightweight simulation
and trace resampling.

@@

The choice of a simulator is a critical part of experimental validation that
raises a tradeoff between precision and runtime. High precision can be obtained
by carefully modeling the platform and its network topology, or extracting
information from the jobs from their sources or post-morted logs. This is the
approach used by high-fidelity batch scheduling simulators such as
Batsim~\cite{batsim}. In the present work, the focus is on studying the
EASY-Backfilling mechanism in itself, without adressing the allocation problem.
Moreover, the experimental protocol used here requires many simulation runs.
Therefore, we set the precision/runtime tradeoff at the point which minimizes
simulation runtimes. We discard all topological information relative to the
platform and use the job processing times of the original workloads. In this
setup, the processors are considered to be undistinguishable from each other,
and jobs can be discontinuously mapped to any available processor on the
system. We develop a lightweight simulator\cite{ocst} and release it to the
community. See the reproducibility paragraph below for more information.  With
this simulator, we are able to replay EASY-FCFS on the CEA-CURIE trace in 33
seconds with a machine equipped with an Intel(R) Core(TM) i5-4310U CPU @
2.00GHz. As a comparison, the Batsim simulator with network
communication modeling disabled and no resource contiguity takes more than 7 hours to
replay the same trace on the same hardware. The complete simulation campaign
presented below is executed on a single Dell PowerEdge R730 in 22 hours,
including input preparation and analysis code.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. This variability can be so extreme
that even a timeframe of a fews years can be insufficient to compare scheduling
algorithms. In our experience inverting some months in the trace at random can
lead to reversal in the results. This motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: For evey week in
a resampled trace, we iterate over every existing user in the original trace.
For each of these users, we choose a week of the original trace uniformly at
random. The jobs from this user in this original week are chosen to be present
in the week being resampled.  This simple mechanism enables us to preserve the
dependency within a week.  Although the system state may not be independent
from one week to an other, we rely on existing work~\cite{jsspp17} and assume
for these experiments that the system is in a stationary state. While we
do provide guidelines as how to incorporate tracking variant of our algorithms
in Section~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
this will not be studied in the experiments from this paper, and is left for
future research.

@@

We use devise the following experimental campaign. For each of the 7 workloads
considered, we use the original data to resample 60 traces, each with a length
of two years. For each resampled trace, we run simulations using the
EASY($P,\Theta$) scheduling heuristic. We fix the value of the job waiting time
threshold $\Theta$ at 40 hours, a sane default. The choice of this default
stems from the analysis developed in ~\cite{jsspp17}. This value is high
enough for queue reorderings to be leveraged while being low enough to provide
the functional requirement of preventing starvation. We simulate
EASY($P,\Theta=40h$) using the following strategies:

@@

  \item The simulation-based ("full feedback") strategy with hyperparameter
    $\Delta \in \left\{ 1 \text{day},1 \text{week} \right\}$. This choice of
    hyperparameter is a consequence of the natural rythm of human activity:
    daily or weekly simulation periods seem to facilitate comparability of
    performance between two periods. Indeed, arbitrary period values would
    needlessly increase the variability of the cost metric.

@@

The experimental campaign consists of 7 traces $\times$ 60 resamples $\times$
(12+2+2+2) algorithms, which equate to 12600 scheduling simulations spanning
two years.  This means a total simulated timespan of 25200 years. Moreover,
both the full-feedback and noisy feedback strategies will also be calling
simulations as part of their procedure. Indeed, these strategies re-simulate
the system 12 times every at every period $\Delta$, which bumps the total
effective simulation timespan to 92400 years. This motivates the use of a
lightweight simulator in the context of this research.

@@

We begin by drawing the attention of the reader to Table~\ref{tab:full}, which
reports the average total waiting time performance improvement over
EASY(FCFS,40h) of all the fixed policies defined in ~\ref{subsec:policies}. Let
us denote the resampled traces of a given workload by $T_k$, where k is the
index of the resampled trace ($1 \le k \le 60$). We also denote the number of
jobs in the resampled trace $T_k$ by $l(T_k)$. We denote by $\bar{w}(T_k,S)$
the cumulative waiting time obtained by a simulation run for a strategy $S$:

@@

The scheduling of parallel jobs on a given HPC platform is a very hard
optimization problem with many uncertain parameters. Even though these
uncertainties could be reduced, determining efficient strategies remains
difficult.

@@

More precisely, we showed that it is worth learning on the scheduling process
itself. Indeed, reordering the submission queues under EASY-BackFilling leads
to considerable gains in performance.

This approach was used in two methods: a method based on a simulator and a
method based on a multi-armed bandit algorithm. The first one provides very
good results, reducing the average waiting times of the baseline FCFS ordering
policy by 11\% to 60\%. The second one is slighly less efficient with an
improvement factor of 8\% to 46\%. However, it is easier to use and cheaper
to run.

There are two directions to extend this work.  First, the question of
the existence of a more efficient selection strategy for bandit feedback for
these systems is not addressed by this work. Is a simulator really necessary?
%@Denis: a essayer.. 1 mois de travail pour tester?
Second, one can help but wonder what performance could be achieved by extending
the search space to arbitrary policies. This means doing a complete study
including lower bounding the cost and studying learning performance using some
function approximator.
%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
As these directions are data-bound, they will require careful application of
trace resampling techniques in order to be successful.
